### 2023-01-09 Manuscript/Draft1.Rmd
paper updates

#read in results tables
YT_results<-read.csv("Yellowtail_results.csv")
Cod_results<-read.csv("Cod_results.csv")
Had_results<-read.csv("Haddock_results.csv")
```

@@

Our goal is to determine indexing approaches and settings that are robust to future environmental conditions and resulting spatial biomass patterns. An underlying assumption in all indexing methods is that individual random samples combine to accurately represent true abundance by a) sampling all strata in which the population exists and b) low enough noise level in the samples to allow for a discernible pattern. This assumption can be questioned given enough noise in the sampling process **cite?** and/or shifting spatial preferences driven by climate change causing a population to move into a previously uninhabited strata. To simulate the impact of noise, we compare indexing estimates after adding noise to our samples versus those using the true sampling values. We simulate the effect of populations moving into new habitat by comparing indexing estimates using samples from all strata inhabited by each species on Georges bank to those that only include a subset of the full spatial domain for each species.[[may need to spell the aspect of reduced spatial domain a bit more because it might be confusing to readers who are expected us to just add areas around the current ones instead of reducing the strata]]

When combining population trends for each species, differing temperature scenarios, altering seasons, and sampling possibilities (noise, strata, covariates) there are a large number of scenario combinations to consider. The columns in Table COMBOS show the choices that define each scenario.

<div align="center">Table COMBOS Each index estimate chooses one condition from each of the following columns. There are $3*3*2*2*2*2*2=288$ VAST model combinations and $3*3*2*2*2=144$ stratified mean estimates.</div>

```{r scenarios, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}

tabl <- "
|	Species	|	Population Trend | Temperature Scenario | Strata Included | Noise Added | Covariates (VAST) | Season
|:--	|:-- |:--- |:-- |:-- |:-- |:-
|Yellowtail | Increasing | Repeating | All strata | No| No Covariates | Spring
|Cod | Constant | Increasing 5$^{\\circ}$| Subset | Yes| Temp + Habitat | Fall
|Haddock | Decreasing | | | | |
"

#knitr::kable(tabl, vline = "|")

cat(tabl) # output the table in a format good for HTML/PDF/docx conversion

@@

## Data and Code Availability
All data and code used in this work are available at https://github.com/cmlegault/IBMWG. [[need to update this]]

@@

param_table1 = read_csv("Parameters1.csv")

@@

param_table = read_csv("Parameters.csv")

@@

result_table = read_csv("Yellowtail_results_final.csv")

@@

param_table1 = read_csv("Cod_results.csv")

>>>>

#read in results tables
YT_results<-read.csv("CSVs/Yellowtail_results.csv")
Cod_results<-read.csv("CSVs/Cod_results.csv")
Had_results<-read.csv("CSVs/Haddock_results.csv")
```

@@

Our goal is to determine indexing approaches and settings that are robust to future environmental conditions and resulting spatial biomass patterns. An underlying assumption in all indexing methods is that individual random samples combine to accurately represent true abundance by a) containing a low enough noise level in the samples to allow for a discernible pattern and b) sampling all strata in which the population exists. These assumptions can be questioned given enough noise in the sampling process **cite?** and/or climate change causing a population to move into previously uninhabited strata. To simulate the impact of noise, we compare indexing estimates after adding noise to our samples versus those using the true sampling values. **BL: Help with correct notation for adding noise**. We simulate the effect of populations moving into new habitat by comparing indexing estimates using samples from all strata inhabited by each species on Georges bank to those that only include a subset of the full spatial domain for each species. We chose strata to exclude for each species by reviewing how spatial preferences evolved in our increasing temperature scenarios and removing strata that each species either shifted into, or away from. Tabl XXX lists all strata inhabited by each species, those that are removed from certain calculations, and the explanation of why these strata were removed.

**BL: I tried to provid emore detail above and also added another table**
[[may need to spell the aspect of reduced spatial domain a bit more because it might be confusing to readers who are expected us to just add areas around the current ones instead of reducing the strata]]

When combining population trends for each species, differing temperature scenarios, altering seasons, and sampling possibilities (noise, strata, covariates) there are a large number of scenario combinations to consider. The columns in Table \@ref(tab:scenarios) show the choices that define each scenario.

```{r scenarios,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}

options(knitr.kable.NA = '')

combos_table = read_csv("CSVs/scenario_combos.csv")

summary_tab <- combos_table %>% 
   mutate_if(grepl('5deg',.), ~replace(., grepl('5deg', .), paste("Increasing 5","\U00B0","C",sep="")))

# Create table to present the data]
 kable(summary_tab, "simple",
                     caption = "Each index estimate chooses one condition from each of the following
225 columns. There are 3∗3∗2∗2∗2∗2∗2 = 288 VAST model combinations and 3∗3∗2∗2∗2 = 144
226 stratified mean estimates.",
          format = "latex", booktabs = TRUE) %>%
          kable_styling(latex_options = "scale_down")
```

@@

## Data and Code Availability
All data and code used in this work are available at https://github.com/Blevy2/READ-PDB-blevy2-MFS2.

@@

param_table1 = read_csv("CSVs/Parameters1.csv")

@@

param_table = read_csv("CSVs/Parameters.csv")

@@

result_table = read_csv("CSVs/Yellowtail_results_final.csv")

@@

param_table1 = read_csv("CSVs/Cod_results.csv")

### 2023-01-09 Manuscript/Draft1.tex
paper updates

Our goal is to determine indexing approaches and settings that are robust to future environmental conditions and resulting spatial biomass patterns. An underlying assumption in all indexing methods is that individual random samples combine to accurately represent true abundance by a) sampling all strata in which the population exists and b) low enough noise level in the samples to allow for a discernible pattern. This assumption can be questioned given enough noise in the sampling process \textbf{cite?} and/or shifting spatial preferences driven by climate change causing a population to move into a previously uninhabited strata. To simulate the impact of noise, we compare indexing estimates after adding noise to our samples versus those using the true sampling values. We simulate the effect of populations moving into new habitat by comparing indexing estimates using samples from all strata inhabited by each species on Georges bank to those that only include a subset of the full spatial domain for each species.{[}{[}may need to spell the aspect of reduced spatial domain a bit more because it might be confusing to readers who are expected us to just add areas around the current ones instead of reducing the strata{]}{]}

When combining population trends for each species, differing temperature scenarios, altering seasons, and sampling possibilities (noise, strata, covariates) there are a large number of scenario combinations to consider. The columns in Table COMBOS show the choices that define each scenario.

Table COMBOS Each index estimate chooses one condition from each of the following columns. There are \(3*3*2*2*2*2*2=288\) VAST model combinations and \(3*3*2*2*2=144\) stratified mean estimates.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.19}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.10}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Population Trend
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Temperature Scenario
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Strata Included
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Noise Added
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Covariates (VAST)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Season
\end{minipage} \\
\midrule
\endhead
Yellowtail & Increasing & Repeating & All strata & No & No Covariates & Spring \\
Cod & Constant & Increasing 5\(^{\circ}\) & Subset & Yes & Temp + Habitat & Fall \\
Haddock & Decreasing & & & & & \\
\bottomrule
\end{longtable}

@@

\caption{\label{tab:unnamed-chunk-1}Haddock error results}
\centering
\begin{tabular}[t]{l|l|l|l|r|r|r|l|l}
\hline
Temp & Strata & Noise & season & VAST.No.Cov & VAST.w..Cov & Stratified.Mean & X & X.1\\
\hline
const & all & no & spring & 0.49 & 0.18 & 0.18 & NA & \\
\hline
const & all & yes & spring & 0.73 & 0.43 & 0.21 & NA & Haddock\\
\hline
const & all & no & fall & 0.28 & 0.05 & 0.26 & NA & Increasing Population\\
\hline
const & all & yes & fall & 0.41 & 0.06 & 0.27 & NA & \\
\hline
const & reduced & no & spring & 0.34 & 0.35 & 0.45 & NA & \\
\hline
const & reduced & yes & spring & 0.30 & 0.33 & 0.46 & NA & \\
\hline
const & reduced & no & fall & 0.36 & 0.48 & 0.54 & NA & \\
\hline
const & reduced & yes & fall & 0.33 & 0.46 & 0.52 & NA & \\
\hline
increasing & all & no & spring & 0.25 & 0.05 & 0.26 & NA & \\
\hline
increasing & all & yes & spring & 0.30 & 0.06 & 0.31 & NA & \\
\hline
increasing & all & no & fall & 0.89 & 0.23 & 0.40 & NA & \\
\hline
increasing & all & yes & fall & 1.04 & 0.35 & 0.42 & NA & \\
\hline
increasing & reduced & no & spring & 0.32 & 0.40 & 0.44 & NA & \\
\hline
increasing & reduced & yes & spring & 0.38 & 0.37 & 0.37 & NA & \\
\hline
increasing & reduced & no & fall & 0.44 & 0.64 & 0.72 & NA & \\
\hline
increasing & reduced & yes & fall & 0.42 & 0.62 & 0.70 & NA & \\
\hline
\end{tabular}
\end{table}

@@

All data and code used in this work are available at \url{https://github.com/cmlegault/IBMWG}. {[}{[}need to update this{]}{]}

@@

 \caption{\label{tab:paramsALL2}Parameters used in all population models.}
 \centering
 \fontsize{10}{12}\selectfont
 \begin{tabular}[t]{lllllll}
 \toprule
 Parameter & Description & Unit & Yellowtail & Cod & Haddock & Source\\
 \midrule
 <U+03C1> & Ford's growth coefficient & 1/wk & 4.48 & 4.43 & 4.49 & NA\\
 M & Natural mortality & 1/wk & 0.2064 & 0.2728 & 0.334 & NA\\
 W1 & Weight of fully recruited fish & kg & 0.39 & 2.95 & 1.12 & NA\\
 W2 & Weight of pre-recruit fish & kg & 0.13 & 0.39 & 0.19 & NA\\
 sigma & Variance in recruited fish & kg*kg & 0.55 & 0.55 & 0.55 & NA\\
 \addlinespace
 lambda & Decay rate for movement & - & 0.7 & 0.7 & 0.7 & NA\\
 spwn & Spawning weeks for species s & wks & 9-12 & 8-13 & 11-14 & NA\\
 rec & Recruitment weeks for species s & wks & 9-12 & 8-13 & 11-14 & NA\\
 \bottomrule
 \end{tabular}
 \end{table}

>>>>

Our goal is to determine indexing approaches and settings that are robust to future environmental conditions and resulting spatial biomass patterns. An underlying assumption in all indexing methods is that individual random samples combine to accurately represent true abundance by a) containing a low enough noise level in the samples to allow for a discernible pattern and b) sampling all strata in which the population exists. These assumptions can be questioned given enough noise in the sampling process \textbf{cite?} and/or climate change causing a population to move into previously uninhabited strata. To simulate the impact of noise, we compare indexing estimates after adding noise to our samples versus those using the true sampling values. \textbf{BL: Help with correct notation for adding noise}. We simulate the effect of populations moving into new habitat by comparing indexing estimates using samples from all strata inhabited by each species on Georges bank to those that only include a subset of the full spatial domain for each species. We chose strata to exclude for each species by reviewing how spatial preferences evolved in our increasing temperature scenarios and removing strata that each species either shifted into, or away from. Tabl XXX lists all strata inhabited by each species, those that are removed from certain calculations, and the explanation of why these strata were removed.

\textbf{BL: I tried to provid emore detail above and also added another table}
{[}{[}may need to spell the aspect of reduced spatial domain a bit more because it might be confusing to readers who are expected us to just add areas around the current ones instead of reducing the strata{]}{]}

When combining population trends for each species, differing temperature scenarios, altering seasons, and sampling possibilities (noise, strata, covariates) there are a large number of scenario combinations to consider. The columns in Table \ref{tab:scenarios} show the choices that define each scenario.

\caption{\label{tab:scenarios}Each index estimate chooses one condition from each of the following
225 columns. There are 3*3*2*2*2*2*2 = 288 VAST model combinations and 3*3*2*2*2 = 144
226 stratified mean estimates.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllllll}
\toprule
Species & Population Trend & Temperature Scenario & Strata Included & Noise Added & Covariates (VAST) & Season\\
\midrule
Yellowtail & Increasing & Repeating & All strata & No & Spring & \\
Cod & Constant & Increasing 5°C & Subset & Yes & Fall & \\
Haddock & Decreasing &  &  &  &  & \\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-1}Haddock error results}
\centering
\begin{tabular}[t]{l|l|l|l|r|r|r|l|l}
\hline
Temp & Strata & Noise & season & VAST.No.Cov & VAST.w..Cov & Stratified.Mean & X & X.1\\
\hline
const & all & no & spring & 0.49 & 0.18 & 0.18 &  & \\
\hline
const & all & yes & spring & 0.73 & 0.43 & 0.21 &  & Haddock\\
\hline
const & all & no & fall & 0.28 & 0.05 & 0.26 &  & Increasing Population\\
\hline
const & all & yes & fall & 0.41 & 0.06 & 0.27 &  & \\
\hline
const & reduced & no & spring & 0.34 & 0.35 & 0.45 &  & \\
\hline
const & reduced & yes & spring & 0.30 & 0.33 & 0.46 &  & \\
\hline
const & reduced & no & fall & 0.36 & 0.48 & 0.54 &  & \\
\hline
const & reduced & yes & fall & 0.33 & 0.46 & 0.52 &  & \\
\hline
increasing & all & no & spring & 0.25 & 0.05 & 0.26 &  & \\
\hline
increasing & all & yes & spring & 0.30 & 0.06 & 0.31 &  & \\
\hline
increasing & all & no & fall & 0.89 & 0.23 & 0.40 &  & \\
\hline
increasing & all & yes & fall & 1.04 & 0.35 & 0.42 &  & \\
\hline
increasing & reduced & no & spring & 0.32 & 0.40 & 0.44 &  & \\
\hline
increasing & reduced & yes & spring & 0.38 & 0.37 & 0.37 &  & \\
\hline
increasing & reduced & no & fall & 0.44 & 0.64 & 0.72 &  & \\
\hline
increasing & reduced & yes & fall & 0.42 & 0.62 & 0.70 &  & \\
\hline
\end{tabular}
\end{table}

@@

All data and code used in this work are available at \url{https://github.com/Blevy2/READ-PDB-blevy2-MFS2}.

@@

 \caption{\label{tab:paramsALL2}Parameters used in all population models.}
 \centering
 \fontsize{10}{12}\selectfont
 \begin{tabular}[t]{lllllll}
 \toprule
 Parameter & Description & Unit & Yellowtail & Cod & Haddock & Source\\
 \midrule
 <U+03C1> & Ford's growth coefficient & 1/wk & 4.48 & 4.43 & 4.49 & \\
 M & Natural mortality & 1/wk & 0.2064 & 0.2728 & 0.334 & \\
 W1 & Weight of fully recruited fish & kg & 0.39 & 2.95 & 1.12 & \\
 W2 & Weight of pre-recruit fish & kg & 0.13 & 0.39 & 0.19 & \\
 sigma & Variance in recruited fish & kg*kg & 0.55 & 0.55 & 0.55 & \\
 \addlinespace
 lambda & Decay rate for movement & - & 0.7 & 0.7 & 0.7 & \\
 spwn & Spawning weeks for species s & wks & 9-12 & 8-13 & 11-14 & \\
 rec & Recruitment weeks for species s & wks & 9-12 & 8-13 & 11-14 & \\
 \bottomrule
 \end{tabular}
 \end{table}
